---
title: "Statistical Inference Project"
author: "JP Dunlap"
date: "July 27, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Part 1: Simulation Exercise

Investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution is simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Set lambda = 0.2 for all of the simulations. Investigate the distribution of averages of 40 exponentials. Conduct 1000 simulations.

```{r intialization}
## Initialize lambda, mu, and sigma

lambda <- 0.20
mu <- 1/lambda
sigma <- 1/lambda

```

### Discussion

The exponential distribution is the probability distribution in which events occur continuously and independently based on a constant rate of change. It descibes the time between events in a Poission process (https://en.wikipedia.org/wiki/Exponential_distribution).

The exponential distribution has a theoretical mean and standard deviation equal to 1/lambda, where lambda is the rate of change in the underlying Poisson process.


## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
